### 2024-04-18 paper/paper.qmd
Added more changes to paper result section

---
title: "Characteristics of Top Songs Has Changed from Pandemic Brain"
subtitle: "An analysis of songs on Billboard's Year-End Hot 100 singles list (2014 to 2023)"
author: Emily Su
thanks: "Code and data are available at: https://github.com/moonsdust/top-songs."
date: today
date-format: long
abstract: "Music often reflects the current climate of society and there is a growing interest in how hit songs as seen on Billboard's Year-End Hot 100 singles has changed after the COVID-19 pandemic. This paper looks at music characteristics such as tempo, song duration, loudness, and modality of songs from Billboard's Year-End Hot 100 singles list from 2014 to 2023 to reveal patterns and relationships to explain the difference between top songs before 2020 and 2020 onwards. The results show that hit songs from 2020 onwards had become on average shorter, quieter, but slightly faster with the melody of songs being in a major key. These results can support the evaluation of the emotional state of different populations and improve treatments such as music therapy, however further investigation is needed on the influence of lyrics on different music characteristics."
format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(ggcorrplot)
library(ggplot2)
library(knitr)
library(modelsummary)
library(rstanarm)

# Packages related to causal model created
library(DiagrammeR)
library(rsvg)
library(magrittr)
library(DiagrammeRsvg)
library(png)

# Read in data 
playlist_analysis_data <- read_parquet("../data/analysis_data/playlists_analysis_data.parquet")

hit_song_model_results <- readRDS("../models/hit_song_model.rds")
```

# Introduction

During the COVID-19 pandemic, viral songs on social media sites such as TikTok were met with thousands of listens on streaming platforms such as Spotify and at times go onto being on lists like the Billboard Year-End Hot 100. These songs often accompany thousands of short clips of people talking about their lives, challenges, etc. Ghaffari et al in their study, "The impact of COVID-19 on online music listening behaviors in light of listeners’ social interactions" noted that other studies have found that people used music as a way to cope and regulate their thoughts and emotions during the pandemic and the lockdowns that came with it [@citecoping]. We can see this translated through the short-form videos where users don't talk and instead have the song play out to convey how they are feeling or thinking. Ghaffari et al's findings are similar to how music therapy works. As defined by the Canadian Association of Music Therapists, music therapy is the use of music to support an individual's health, social development, and well-being and some of the techniques used include listening to music [@citemusictherapy]. Hurwitz and Krumhansl conducted a study into how people's listening habits throughout the pandemic and they found that the songs they listened to often were linked to an emotion such as sadness and/or invoking memories from the past [@citeshiftingniches]. However, this raises the following question, which we explored in our analysis: how are top songs prior to 2020 or the start of the pandemic different from the top songs during and after 2020? In our analysis, we defined hit songs before 2020 to be hit songs before the pandemic and hit songs during 2020 and afterwards to be hit songs during and after the pandemic.

In this paper, to investigate patterns and trends in music, we analyzed data from the Billboard Year-End Hot 100 singles list from 2014 to 2023 on music characteristics such as its modality (major or minor key), track duration, loudness, and tempo. Our estimand is the song characteristics from the Billboard Year-End Hot 100 singles such as its modality (major or minor), track duration, loudness, and tempo if a song was created before 2020 or during and after 2020. The list allowed us to gauge how preference in music characteristics has changed overall during and after the pandemic. Current studies give us a sense of how currently people are feeling based on surveys done. However, there is currently a lack of understanding and focus on the difference in musical characteristics of popular songs with the general population before the pandemic and during and after it. In our findings, our data showed that songs in a major key made up the majority of top songs before, during, and after 2020 and top songs were on average quieter, shorter, and had slightly faster tempos during and after 2020. [TODO: ADD RESULTS FROM MODEL]. Music can reflect the climate of society and understanding trends in music characteristics of hit songs can help provide insight on the overall feelings of different populations during and after the pandemic as well as before it. This can aid in improving treatments such as music therapy.

In the rest of this paper, the data section (@sec-data) will cover the dataset used, how it was obtained, define the variables of interest that are used by our tables and graphs, and briefly explain the data cleaning process. The model section (@sec-model) will explain our proposed causal model, which would explain potential relationships in our data, the setup of our model to understand these relationships, and justification for our model. In the results section (@sec-results), we will reveal tables and graphs made on our datasets, explain what they show, and show our results from our model. In the discussion section (@sec-discussion), we will connect back to the real world and explain what the results could mean, the implications of our results, potential areas of improvement for the paper, and suggestions for future works. Finally, the appendix section (@sec-appendix) will extra tables and graphs from our results as well, as additional information about the model, and a link to a Shiny application featuring interactive graphs of some of the results.

# Data {#sec-data}

## Methodology
The dataset used in this paper was retrieved, simulated, cleaned, analyzed, and tested using the R programming language [@citeR], tidyverse [@citetidyverse], knitr [@citeknitr], janitor [@citejanitor], dplyr [@citedplyr], ggplot2 [@citeggplot2], spotifyr [@citespotifyr], usethis [@citeusethis], arrow [@citearrow], ggcorrplot [@citeggcorrplot], and testthat [@citetestthat]. Code written in the scripts was checked and styled with lintr [@citelintr] and styler [@citestyler]. Accompanying the paper, a Shiny app [@citeshiny] using shinyWidgets [@citeshinyWidgets] was created containing interactive visualizations. The packages that were used for the model-related sections or used for the model itself are DiagrammeR [@citeDiagrammeR], rsvg [@citersvg], magrittr [@citemagrittr], DiagrammeRsvg [@citeDiagrammeRsvg], png [@citepng], rstanarm [@citerstanarm], and modelsummary [@citemodelsummary].

## Data Source and Measurements
The data used for the model and analysis comes from the ```playlists_analysis_data``` dataset. This dataset was constructed by using an R wrapper for Spotify's API called spotifyr [@citespotifyr] to scrape playlists complied by Spotify users antoniomendoza and wickeddreamer96 of songs from the Billboard Year-End Hot 100 singles from 2014 to 2023 [@citebillboard]. User antoniomendoza created playlists for the years 2018 to 2023, inclusive, while user wickeddreamer96 created playlists for the years 2014 to 2017, inclusive. The latest version of the data used was scraped on Tuesday March 26, 2024 at 9:46pm EDT. The Billboard Year-End Hot 100 singles list for the year 2014 onward was created based on data from Nielsen SoundScan [@citenielsen]. Nielsen SoundScan
collect data weekly from 39,000 retail outlets, venues, etc. in 19 countries such as Canada and the United States on in-store sales and digital sales [@citenielsen]. They also track how often songs are streamed online and played on the radio [@citenielsen]. The list comprise of songs they collected data on starting from December of the previous year to the end of November of the current year, which is when Billboard releases the Year-End Hot 100 list [@citebillboardcalculations].

Initially, we had scraped playlists created by Spotify that they have compiled of the top songs on the app. However, we had an insufficient amount of observations from these playlists for our dataset. The playlists we previously were using only had around 50 songs each and there were years where Spotify had not created a top song playlist for such as 2021 and 2022.

Each observation in the dataset belongs to a song from the Spotify playlist that we scraped compiled by antoniomendoza and wickeddreamer96 for songs on the Billboard Year-End Hot 100 singles. For the duplicate songs that showed up in different years, they were kept as we wanted to analyze song characteristics. In total, there are 1000 observations in the ```playlists_analysis_data``` dataset, which equals to the number of songs on the Billboard Year-End Hot 100 singles lists from 2014 to 2023.
```{r}
#| label: tbl-dataset-preview
#| tbl-cap: Preview of dataset on Billboard Year-End Hot 100 singles from 2014 to 2023 provided by Spotify
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  select(hit_year, track_duration_ms, tempo, loudness, mode_name, key_mode, period) |>
  head() |>
  kable()
```
Not seen in @tbl-dataset-preview, before_pandemic, major, and minor are numerical versions of the mode_name and period columns. The variables that were originally the raw dataset that was scraped are track_duration_ms, tempo, loudness, mode_name, and key_mode. track_duration_ms describes the duration of the song in milliseconds (ms), tempo is the mean (average) beats per minutes (BPM) of the song, and loudness is mean loudness of a top song in decibels (dB). mode_name represents the key the melody of the song is in, which is either "major" or "minor" and key_mode signifies the scale the song is in.

The variables constructed from the raw dataset are before_pandemic, period, hit_year, major, and minor. How these variables were constructed was that hit_year was obtained from check the name of the playlist and it represents the hit. From hit_year, the period column was constructed, which indicates whatever or not the hit year of the song was before 2020, which was when the COVID-19 pandemic started, or during 2020 onwards. Reading from the period column, the before_pandemic column was created, where 1 represents that the song was a hit song before the pandemic and 0 otherwise. The major and minor columns represents whatever or not the melody of the song is in a major or minor key and the value would be 1 if it's that key and 0 otherwise. These columns were constructed from checking mode_name when it would indicate either "major" or "minor".

## Variables of Interest
As seen in @tbl-dataset-preview, there are multiple columns however our main variables of interests are before_pandemic, track_duration_ms, loudness, tempo, and mode_name for our model. The major and minor columns were used initially to visualize potential relationships between variables as seen in @fig-correlation-matrix. period, hit_year, and key_mode were used to visualize our data prior to training our model and to understand the data.

# Model {#sec-model}
From our analysis of the data, we observed that there were trends between a song's duration, loudness, tempo, and modality for top songs prior 2020 and 2020 onwards. We are interested in investigating if a song is likely to be a top song prior to 2020 or not based on what we know about the duration, loudness, tempo, and modality of a top song. We will do this by creating a model.

## Model set-up
In order to predict a top song was a hit prior to 2020, we will make the assumption that there is a relationship between the outcome variable, if the song is a top song prior to 2020 and our variables of interests: duration, loudness, tempo, and modality. Our variables of interests will be the predictor variables for the model. We will define our model as a logistic regression model with a Bayesian approach as follows:
\begin{align*} 
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times \mbox{duration}_i + \beta_2 \times \mbox{loudness}_i + \beta_3 \times \mbox{tempo}_i + \beta_4 \times \mbox{modality}_i\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
\beta_4 &\sim \mbox{Normal}(0, 2.5) \\
\end{align*}

We define $y_i$ to be a top song, which if it is 1 represents a top song prior to 2020 and 0 if it is a top song from 2020 onwards. $\pi_i$ is the probability that a top song $i$ is a top song prior to 2020. Let $\mbox{duration}_i$ be the duration of the top song in milliseconds (ms) and $\mbox{loudness}_i$ to be the mean loudness of a top song in decibels (dB). Set $\mbox{tempo}_i$ to be the mean beats per minutes (BPM) of a top song and $\mbox{modality}_i$ to be the modality of the top song, where 1 means the melody of the song is in a minor key and 0 if it is in a major key.

We ran the model using the rstanarm package [@citerstanarm] and R [@citeR] with the default priors from the rstanarm package. Diagnostics related to the model such as a trace plot, Rhat plot, and posterior predictive check can be found at @sec-appendix.

### Model justification
```{r}
#| label: fig-causal-model
#| fig-cap: Causal relationship between song characteristics and hit song (before 2020 or during and after 2020)
#| echo: false
#| warning: false
#| message: false

# Referenced https://www.erikigelstrom.com/articles/causal-graphs-in-r-with-diagrammer/
# https://stackoverflow.com/questions/42737645/how-to-export-images-of-diagrammer-in-r
causal_model <- "digraph {
  graph []
  # Nodes 
  node [shape = plaintext]
    A [label = 'Hit song before 2020 or during and after 2020']
    B [label = 'Modality']
    C [label = 'Loudness']
    D [label = 'Duration']
    E [label = 'Tempo']
    F [label = 'Lyrics (Confounder)']
  # Edges for non-confounding variables
  edge []
    B->A
    C->A
    D->A
    E->A
{ rank = same; }
  # Edges for confounding variable
  edge [style = dashed]
    F -> B
    F -> C
    F -> D
    F -> E
    F -> A
}"

exporting_image <- grViz(causal_model) |>
  export_svg() |>
  charToRaw() |>
  rsvg_png("../other/sketches/graph_2_casual_model.png", width = 1000)

# NOTE: Had to save diagram and read it in as a image since grViz can be only render in HTML and 
# not in PDFs 
# Referenced https://stackoverflow.com/questions/23861000/displaying-images-in-r-in-version-3-1-0
the_image <- readPNG("../other/sketches/graph_2_casual_model.png")
plot.new()
rasterImage(the_image, 0, 0, 0.8, 1.13)
```
A logistic regression model was chosen because our outcome variable, if the song is a top song prior to 2020, is a binary outcome. However, alternative models were considered. Multiple linear regression model was a candidate for our model since the model has multiple predictor variables but it assumes that the outcome is continuous instead of discrete. Other regression models like negative binomial regression were not chosen since we are not dealing with discrete counts for the outcome but instead discrete binary values of 1 and 0. Our logistic regression model describes the following causal relationship (@fig-causal-model) where song characteristics predicts whatever a hit song was made prior to 2020, excluding a confounding variable, lyrics. The assumption we make is that there is a relationship between song characteristics and a top song being a hit prior to 2020. In @fig-causal-model, the reason lyrics, if there are any, is a confounding variable is because lyrics can affect not only if the song was a hit prior to 2020 but also song characteristics from the way the lyrics are performed. However, we are unable to observe the effects of lyrics directly with our data. Weakness and further improvements to the model are discussed in @sec-discussion.

# Results {#sec-results}
In this section, we visualized our data through graphs and tables as well as present the results from our model.

## Difference in song characteristics before 2020 (the pandemic) and during and after 2020 (the pandemic) of songs from the Billboard Year-End Hot 100 singles

### Scale and Modality
```{r}
#| label: fig-mode-bar
#| fig-cap: "Proportion of songs whose modality is in a major or minor key before 2020 (the pandemic) versus during and after 2020 (the pandemic) from Billboard's Year-End Hot 100 singles"
#| echo: false
#| warning: false
#| message: false

proportion_of_each_mode <- 
  playlist_analysis_data |>
  # Group by if song was a hit before or during and after the pandemic
  group_by(period) |>
  # Count each scale
  count(mode_name) |>
  # rename count column to num_of_key_mode
  rename(
    'num_of_mode_name' = n
  ) |>
  mutate( 
    proportion_mode_name =  round(num_of_mode_name / sum(num_of_mode_name), 2) 
  )

proportion_of_each_mode |> 
  ggplot(mapping = aes(x = mode_name, y = proportion_mode_name)) +
  facet_wrap(facets = vars(period), dir = "h") + 
  geom_bar(stat = "identity", binwidth = 3, fill="#e85569") +
  theme_minimal() +
  labs(x = "Modality",
       y = "Proportion") + 
  theme(axis.text.x = element_text(angle = 90))
```

```{r}
#| label: tbl-mode-table
#| tbl-cap: "Proportion of songs whose modality is in a major or minor key before 2020 (the pandemic) versus during and after 2020 (the pandemic) from Billboard's Year-End Hot 100 singles"
#| echo: false
#| warning: false
#| message: false

proportion_of_each_mode |> 
  rename( 
    "Period" = period,
    "Modality" = mode_name,
    "Count of each mode" = num_of_mode_name,
    "Proportion of each mode" = proportion_mode_name
    ) |>
  kable()
```
@fig-mode-bar and @tbl-mode-table reveals that a proportion of hit songs before and during and after 2020 that had melodies in a major key was around ~60% (61% before 2020 and 63% from 2020 onwards). Songs in a minor key made up about 40% or less of songs (39% before 2020 and 37% from 2020 onwards) from 2014 to 2023 on Billboard's Year-End Hot 100 singles list. There was a slight increase in songs in a major key from 2020 afterwards however songs in a major key making up a majority of Billboard's Year-End Hot 100 singles stayed consist before 2020 and 2020 onwards.

```{r}
#| label: fig-key-mode-bar
#| fig-cap: "Proportion of songs in different scales before 2020 (the pandemic) versus during and after 2020 (the pandemic) from Billboard's Year-End Hot 100 singles"
#| echo: false
#| warning: false
#| message: false

proportion_of_each_key_mode <- 
  playlist_analysis_data |>
  # Group by if song was a hit before or during and after the pandemic
  group_by(period) |>
  # Count each scale
  count(key_mode) |>
  # rename count column to num_of_key_mode
  rename(
    'num_of_key_mode' = n
  ) |>
  mutate( 
    proportion_key_mode =  round(num_of_key_mode / sum(num_of_key_mode), 2)
  )

proportion_of_each_key_mode |> 
  ggplot(mapping = aes(x = key_mode, y = proportion_key_mode)) +
  facet_wrap(facets = vars(period), dir = "v") + 
  geom_bar(stat="identity", fill="#3E509A") +
  theme_minimal() +
  labs(x = "Scale",
       y = "Proportion") + 
  theme(axis.text.x = element_text(angle = 90))
```
When looking into the specific scales of the songs, @fig-key-mode-bar and @tbl-scale-table shows that there was an increase in songs in C major during and after 2020 with C major and C# major made up a majority of songs from Billboard's list from 2020 onwards at approximately 9% each. Prior to 2020, the proportion of songs in C major were at 11% however there was a decline in songs in C major after the start of the pandemic. There was a decline in songs in G# major from 8% of the list to 6% of the list as well as a an increase in songs in D minor and D# minor from less than 1% of the list to 2% each. There were also a decline in songs in F minor where they would make up 6% of Billboard's list prior to 2020 to 4% during and after 2020.

\newpage

### Track Duration
```{r}
#| label: fig-track-duration-scatterplot
#| fig-cap: "Relationship between the hit year and track duration of a song (in ms) on Billboard's Year-End Hot 100 singles list. Songs from the same year are grouped in the same colour and the horizontal, dashed line divides the hit tracks before 2020 and hit tracks from 2020 and beyond. Blue, smoother line describes the overall trend and relationship between hit year and track duration."
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  ggplot(aes(x = hit_year, y = track_duration_ms, color = hit_year)) +
  geom_smooth() +
  geom_jitter() +  # So we can see all points
  geom_point(alpha = 0.3) +
  theme_minimal() +
  labs(x = "Hit year",
       y = "Duration of song (in ms)") + 
  geom_vline(xintercept = 2020, linetype = "dashed") +
  scale_color_gradient(low = "#3E509A", high = "#e85569", guide = "none")
```

```{r}
#| label: fig-track-duration-hist
#| fig-cap: "Distribution of track duration (in ms) before 2020 (the pandemic) versus during and after 2020 (the pandemic) from songs on Billboard's Year-End Hot 100 singles list."
#| echo: false
#| warning: false
#| message: false
#| 
playlist_analysis_data |> 
  ggplot(mapping = aes(x = track_duration_ms)) +
  facet_wrap(facets = vars(period), dir = "v") + 
  geom_histogram(binwidth = 7000, fill="#e85569") +
  theme_minimal() +
  labs(x = "Duration of song (in ms)",
       y = "Count")
```

```{r}
#| label: fig-track-duration-boxplot
#| fig-cap: "Track duration (in ms) before 2020 (the pandemic) versus during and after 2020 (the pandemic) from songs on Billboard's Year-End Hot 100 singles list."
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  ggplot(mapping = aes(x = track_duration_ms)) +
  facet_wrap(facets = vars(period), dir = "v") + 
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Duration of song (in ms)",
       y = "Count")
```
```{r}
#| label: tbl-track-duration-stats-before
#| tbl-cap: "Minimum, quartiles, median, and maximum of track duration (in ms) before 2020 (the pandemic) on Billboard's Year-End Hot 100 list"
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  # Filter for "Before Pandemic"
  filter(period == "Before Pandemic") |>
  # Choose track_duration_ms
  select(track_duration_ms) |>
  # Rename columns
  rename (
    "Duration of song (in ms)"= track_duration_ms
  ) |>
  summary() |>
  kable()
```

```{r}
#| label: tbl-track-duration-stats-during-after
#| tbl-cap: "Minimum, quartiles, median, and maximum of track duration (in ms) during and after 2020 (the pandemic) on Billboard's Year-End Hot 100 list"
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  # Filter for During and After Pandemic
  filter(period == "During and After Pandemic") |>
  # Choose track_duration_ms
  select(track_duration_ms) |>
  # Rename columns
  rename (
    "Duration of song (in ms)"= track_duration_ms
  ) |>
  summary() |>
  kable()
```
@fig-track-duration-scatterplot reveals that songs over the years from 2014 to 2020 decreased slightly in duration and from 2020 onwards, there was almost no change in trend overall based on the blue, smoother line. @fig-track-duration-hist shows that the distribution for top songs before the pandemic and during and after the pandemic are skewed to the right. According to  @tbl-track-duration-stats-during-after and @tbl-track-duration-stats-before, the distribution of top songs during and after the pandemic are more disperse around their mean of 195999 ms than those from before the pandemic where there is less variance, which means how close data points are to their mean/average value. Based on the distribution, there are more songs with duration of about 216286 ms for top songs before the pandemic. @fig-track-duration-boxplot, @tbl-track-duration-stats-during-after, and @tbl-track-duration-stats-before also show that hit songs during and after the pandemic are shorter in length on average with a mean track duration of 195999 ms compared to hit songs from before the the pandemic with a mean track duration of 216286 ms. The median (the middle value of the data when sorted) track duration also decreased from 214070 ms before 2020 to 195120 ms for 2020 onwards.

### Loudness
```{r}
#| label: fig-loudness-scatterplot
#| fig-cap: "Relationship between the hit year and loudness of a song (in dB) on Billboard's Year-End Hot 100 singles list. Songs from the same year are grouped in the same colour and the horizontal, dashed line divides the hit tracks before 2020 and hit tracks from 2020 and beyond. Blue, smoother line describes the overall trend and relationship between hit year and loudness." 
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  ggplot(aes(x = hit_year, y = loudness, color = hit_year)) +
  geom_smooth() +
  geom_jitter() +  # So we can see all points
  geom_point(alpha = 0.3) +
  theme_minimal() +
  labs(x = "Hit year",
       y = "Loudness of song (in dB)") + 
  geom_vline(xintercept = 2020, linetype = "dashed") +
  scale_color_gradient(low = "#3E509A", high = "#e85569", guide = "none")
```
```{r}
#| label: fig-loudness-hist
#| fig-cap: "Distribution of loudness (in dB) of hit songs before 2020 (the pandemic) versus during and after 2020 (the pandemic) from Billboard's Year-End Hot 100 list"
#| echo: false
#| warning: false
#| message: false
playlist_analysis_data |> 
  ggplot(mapping = aes(x = loudness)) +
  facet_wrap(facets = vars(period), dir = "v") + 
  geom_histogram(binwidth = 0.3, fill="#3E509A") +
  theme_minimal() +
  labs(x = "Loudness of song (in dB)",
       y = "Count")
```
```{r}
#| label: tbl-loudness-stats-before
#| tbl-cap: "Minimum, quartiles, median, and maximum of loudness of song (in dB) before 2020 (the pandemic) on Billboard's Year-End Hot 100 list"
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  # Filter for "Before Pandemic"
  filter(period == "Before Pandemic") |>
  # Choose loudness
  select(loudness) |>
  # Rename columns
  rename (
    "Loudness of song (in dB)"= loudness
  ) |>
  summary() |>
  kable()
```
```{r}
#| label: tbl-loudness-stats-during-after
#| tbl-cap: "Minimum, quartiles, median, and maximum of loudness of song (in dB) during and after 2020 (the pandemic) on Billboard's Year-End Hot 100 list"
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  # Filter for "During and After Pandemic"
  filter(period == "During and After Pandemic") |>
  # Choose loudness
  select(loudness) |>
  # Rename columns
  rename (
    "Loudness of song (in dB)"= loudness
  ) |>
  summary() |>
  kable()
```
@fig-loudness-scatterplot shows that from 2014 to 2023, there was a decrease in loudness overall with hit songs as see with the negative slope of the blue, smoother line. @fig-loudness-hist reveals that the distribution of top songs' loudness is left-skewed with the loudness of top songs from during and after the pandemic being more disperse than those before the pandemic around their mean. This suggests that top songs from 2020 onwards varied more in terms of their loudness when compared with top songs before 2020. @tbl-loudness-stats-before and @tbl-loudness-stats-during-after shows that before the pandemic, top songs on Billboard's list had a mean loudness of -5.964 dB while during and after the pandemic, it had a mean loudness of -6.516 dB. The median loudness of top tracks has also decrease from -5.700 dB prior to the pandemic to -6.151 dB during and after the pandemic. This indicates that hit songs during and after the pandemic are less louder than hit songs before the pandemic.

### Tempo
```{r}
#| label: fig-tempo-scatterplot
#| fig-cap: "Relationship between the hit year and tempo of a song (the beats per minutes (BPM) of a song) on Billboard's Year-End Hot 100 singles list. Songs from the same year are grouped in the same colour and the horizontal, dashed line divides the hit tracks before 2020 and hit tracks from 2020 and beyond. Blue, smoother line describes the overall trend and relationship between hit year and tempo." 
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  ggplot(aes(x = hit_year, y = tempo, color = hit_year)) +
  geom_smooth() +
  geom_jitter() +  # So we can see all points
  geom_point(alpha = 0.3) +
  theme_minimal() +
  labs(x = "Hit year",
       y = "Tempo of song (BPM)") + 
  geom_vline(xintercept = 2020, linetype = "dashed") +
  scale_color_gradient(low = "#3E509A", high = "#e85569", guide = "none")
```
```{r}
#| label: fig-tempo-hist
#| fig-cap: "Distribution of tempo (BPM) before 2020 (the pandemic) versus during and after 2020 (the pandemic) from Billboard's Year-End Hot 100 list"
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  ggplot(mapping = aes(x = tempo)) +
  facet_wrap(facets = vars(period), dir = "v") + 
  geom_histogram(binwidth = 3, fill="#e85569") +
  theme_minimal() +
  labs(x = "Tempo of song (BPM)",
       y = "Count")
```
```{r}
#| label: tbl-tempo-stats-before
#| tbl-cap: "Minimum, quartiles, median, and maximum of tempo of a song (BPM) before 2020 (the pandemic) from Billboard's Year-End Hot 100 list"
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  # Filter for "Before Pandemic"
  filter(period == "Before Pandemic") |>
  # Choose tempo
  select(tempo) |>
  # Rename columns
  rename (
    "Tempo of song (BPM)"= tempo
  ) |>
  summary() |>
  kable()
```
```{r}
#| label: tbl-tempo-stats-during-after
#| tbl-cap: "Minimum, quartiles, median, and maximum of tempo of a song (BPM) during and after 2020 (the pandemic) from Billboard's Year-End Hot 100 list"
#| echo: false
#| warning: false
#| message: false

playlist_analysis_data |> 
  # Filter for "During and After Pandemic"
  filter(period == "During and After Pandemic") |>
  # Choose tempo
  select(tempo) |>
  # Rename columns
  rename (
    "Tempo of song (BPM)"= tempo
  ) |>
  summary() |>
  kable()
```
@fig-tempo-scatterplot reveals almost no change in tempo overall across top songs from 2014 to 2023 based on the blue smoother line appearing to be flat. However, looking at @tbl-tempo-stats-before and @tbl-tempo-stats-during-after it shows that both the median and mean tempo of top songs before 2020 (about 120 BPM and 121 BPM, respectively) are lower or about equal than the median and mean tempo of top songs during and after 2020, which is about 120 BPM and 122 BPM, respectively. @fig-tempo-hist shows that in both periods of time, their distributions have skewness close to zero however there is less dispersion around the mean for hit songs prior to 2020. These results suggest that the the tempo of top songs on Billboard's list are slightly faster during and after 2020.

\newpage
## Model Results

@tbl-model-result and @fig-model-result-ci shows the results of our logistic regression model. @tbl-model-result shows the coefficient estimates obtained for our model with their mean absolute deviation (MAD) while @fig-model-result-ci showcases the coefficient estimates (the dot) alongside their 90% credible interval (the line). Coefficient estimates represents....[@citecoefficientsdef]. Mean absolute deviation is....

@tbl-model-result indicates that the y-intercept is -2.002, this however has no meaning in this situation since we only have two outcomes, 0 and 1 to represent a hit song from 2020 onwards and a hit song prior to 2020, respectively. @tbl-model-result and @fig-model-result-ci reveal that the coefficient of track duration is 0.000 when tempo, mode, and loudness remains constant and its credible interval....... [The credible interval means.......] This indicates there is no relationship between a song being a hit before 2020 and the duration of the song.

```{r}
#| label: tbl-model-result
#| tbl-cap: Relationships of a hit song is a hit before 2020 / the pandemic with its tempo, song duration, loudness, and modality with their mean absolute deviation (MAD) values in parenthesis. 
#| echo: false
#| warning: false
#| message: false

modelsummary(
  list("A hit song before 2020" = hit_song_model_results),
  statistic = "mad"
)
```

```{r}
#| label: fig-model-result-ci
#| fig-cap: The credible intervals (line) for coefficient estimates (dot) of predictors of a hit song before 2020
#| echo: false
#| warning: false
#| message: false

modelplot(hit_song_model_results, conf_level = 0.9) +
  labs(x = "90 per cent credible interval")
```

\newpage

# Discussion {#sec-discussion}

[To Do]

Weakness: There were more songs available prior to 2020 in our dataset. This could have impacted the model.

\newpage

\appendix

# Appendix {#sec-appendix}

## Shiny application 
A Shiny application featuring interactive graphs from the paper can be found here:
https://49z7k8-emily-su.shinyapps.io/top-songs-characteristics-app/

## Posterior predictive check for model {#sec-model-details-posterior-check}
@fig-posteriorpredictive shows that the posterior distribution of the fitted model is consistent with the actual data. 
```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-posteriorpredictive
#| fig-cap: "How the data impacts how the model fits"
#| fig-subcap: ["Posterior prediction check for the model"]
#| layout-ncol: 1

rstanarm::pp_check(hit_song_model_results) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Diagnostics for model {#sec-model-details-diagnostics}
@fig-convergencecheck shows that the sampling algorithm used, the Markov chain Monte Carlo (MCMC) algorithm, did not run into issues as the posterior distribution for the model was created. Both graphs do not show anything abnormal.
```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-convergencecheck
#| fig-cap: "Checking the convergence of the Markov chain Monte Carlo (MCMC) algorithm for hit song model"
#| fig-subcap: ["Trace plot for model", "Rhat plot for model"]
#| layout-ncol: 2

plot(hit_song_model_results, "trace")

plot(hit_song_model_results, "rhat") + theme_minimal()
```
\newpage
## Additional Figures

```{r}
#| label: fig-correlation-matrix
#| fig-cap: Correlation between numerical characteristics of songs from Billboard Year-End Hot 100 singles list and if song was a hit before 2020/the pandemic
#| echo: false
#| warning: false
#| message: false

# Referenced https://rpubs.com/Alema/1000474
numeric_playlist_analysis_data <- 
  playlist_analysis_data |>
  select(-c(mode_name, key_mode, period, hit_year))

ggcorrplot(cor(numeric_playlist_analysis_data), 
           hc.order = TRUE,
           outline.color = "white", lab = TRUE, legend.title = "Correlation") 
```

\newpage

## Additional Tables

```{r}
#| label: tbl-scale-table
#| tbl-cap: "Proportion of songs in different scales before 2020 (the pandemic) versus during and after 2020 (the pandemic) from from Billboard's Year-End Hot 100 singles"
#| echo: false
#| warning: false
#| message: false

proportion_of_each_key_mode |> 
  rename( 
    "Period" = period,
    "Scale" = key_mode,
    "Count of each scale" = num_of_key_mode,
    "Proportion of each scale" = proportion_key_mode
    ) |>
  kable()
```

\newpage

# References